\thetitle

### 2021-05-23 writing/rational-natural.tex
TeX setting

When the buggy untyped module of a program is replaced by the typed counterpart,
the type checker fails because this module causes the impedance mismatch. Hence, a
trail that ends at an ill-typed scenario successfully pinpoints the location of
the bug. 
\begin{quote}
\it A Natural blame trail $\conf_0,...\conf_n$ in a lattice $\lattice{\system}$ is
\emph{successful} iff (the program for) its last scenario $\conf_n$ does not type check.  A Natural
blame trail $\conf_0,..,\conf_n$ in a lattice $\lattice{\system}$ is \emph{failing}
iff (the program for) $\conf_n$ type checks and the trail cannot be extended further.
\end{quote}
That is, failing Natural blame trails are those that end in a scenario that does not reveal the bug statically, yet also does not blame
an untyped module. Thus the rational programmer has
no further hints on how to continue the search for the bug.

@@

The definition of success for a Natural exception trail follows that for
a Natural blame trail.
Together, the definitions for the two modes allow the comparison of the usefulness of blame 
with that of mere exceptions for debugging a scenario in the context of Natural semantics.
\begin{quote}
\it 
  Given a program $\system$ and a root $\conf_0$ in $\lattice{\system}$,
  Natural blame is \emph{more useful} than Natural exceptions for
  debugging $\conf_0$ iff 
  the Natural blame trail 
  that starts at $\conf_0$ is successful while the Natural exception trail that
  starts at $\conf_0$ is failing.
\end{quote}

>>>>

When the buggy untyped module of a program is replaced by the typed counterpart,
the type checker fails because this module causes the impedance mismatch. Hence, a
trail that ends at an ill-typed scenario successfully pinpoints the location of
the bug. 
\begin{quote}
\it A {\em Natural blame trail\/} $\conf_0,...\conf_n$ in a lattice $\lattice{\system}$ is
\emph{successful}\/ iff (the program for) its last scenario $\conf_n$ does not type check.  A Natural
blame trail $\conf_0,..,\conf_n$ in a lattice $\lattice{\system}$ is \emph{failing\/}
iff (the program for) $\conf_n$ type checks and the trail cannot be extended further.
\end{quote}
That is, failing Natural blame trails are those that end in a scenario that does not reveal the bug statically, yet also does not blame
an untyped module. Thus the rational programmer has
no further hints on how to continue the search for the bug.

@@

The definition of success for a Natural exception trail follows that for
a Natural blame trail.
Together, the definitions for the two modes allow the comparison of the usefulness of blame 
with that of mere exceptions for debugging a scenario in the context of Natural semantics.
\begin{quote}
\it 
  Given a program $\system$ and a root $\conf_0$ in $\lattice{\system}$,
  Natural blame is \emph{more useful\/} than Natural exceptions for
  debugging $\conf_0$ iff 
  the Natural blame trail 
  that starts at $\conf_0$ is successful while the Natural exception trail that
  starts at $\conf_0$ is failing.
\end{quote}

### 2021-05-23 writing/rational-transient.tex
TeX setting

\begin{quote}
\it A \emph{Transient-first blame trail}/\emph{Transient-last blame trail} is a sequence of scenarios
$\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
$\conf_i \subset \conf_{i+1}$ and
\[ \conf_{i+1} \setminus \conf_i = %
\left\{ %
\begin{tabular}{ll} %
$\{\first{\mblame{\system, \conf_i}}\}$  & if $\conf_i$ produces blame \\ %
$\{\exception{\system, \conf_i}\}$     & otherwise \\ %
\end{tabular} \right. %
\]
where $\first{\mblame{\system, \conf}}$/$\last{\mblame{\system, \conf}}$ is the
first/last module, respectively,  that Transient adds to the blame sequence for $\conf$.
\end{quote}

>>>>

\begin{quote}
\it A \emph{Transient-first blame trail}{\lrdef}\emph{Transient-last blame trail} is a sequence of scenarios
$\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
$\conf_i \subset \conf_{i+1}$ and
\[ \conf_{i+1} \setminus \conf_i = %
\left\{ %
\begin{tabular}{ll} %
$\{\first{\mblame{\system, \conf_i}}\}$  & if $\conf_i$ produces blame \\ %
$\{\exception{\system, \conf_i}\}$     & otherwise \\ %
\end{tabular} \right. %
\]
where $\first{\mblame{\system, \conf}}${\lrdef}$\last{\mblame{\system, \conf}}$ is the
first{\lrdef}last module, respectively,  that Transient adds to the blame sequence for $\conf$.
\end{quote}
